---
title: "Analysis of the Palmer Penguins Dataset"
author: "Moaz Mohammed El-Esawey"
date: "10/25/2021"
output:
  html_document: default
  pdf_document: default
---

## Loading the dataset and packages used
we load the *palmerpenguins* package as it contains the data for out analysis. and tidyverse which contains some packages like *dplyr*, *ggplot* and more that is useful for analysis
```{r echo=TRUE, warning=FALSE, message=FALSE}
library('palmerpenguins')
library('tidyverse')
```


```{r echo=TRUE}
## clean the data and remove all the NA values.
penguins <- penguins %>% 
  drop_na()

dim(penguins)
head(penguins)

```

```{r}
unique(penguins$species)
unique(penguins$island)
```

as we saw here the data contains a 344 records of penguins of different species **Adelie**, **Gentoo** and **Chinstrap** in 3 different islands *Torgersen*, *Biscoe* and *Dream*

the data have recorded is **bill_length_mm**, **bill_depth_mm**, **flipper_length_mm** and **body_mass_g** for each penguin.

## Summary statictics for the dataset
```{r}
penguins %>% 
  group_by(species) %>% 
  drop_na() %>% 
  summarise(mean_fl=mean(flipper_length_mm), mean_bm=mean(body_mass_g)) %>% 
  arrange(mean_fl)
```

as we show above the **Gentoo** specie is the biggest flipper_length_mm and body_mass_g we means that it's the largest penguin in all species.

## Visualizations
```{r}
ggplot(data=penguins, aes(x=flipper_length_mm, y=body_mass_g))+
  geom_point(aes(color=species))
```

As we see that the **Gentoo** specie is the biggest and also found to be alone and with its own cluster of points away from the other two species.

```{r}
ggplot(data=penguins, aes(x=flipper_length_mm, y=body_mass_g))+
  geom_point(aes(color=species, shape=island))
```

Also I found that the **Gentoo** specie found to live alone from the other in the **Biscoe** island.
```{r}
ggplot(data=penguins, aes(x=bill_length_mm, y=body_mass_g))+
  geom_point(aes(color=species))+
  geom_smooth(aes(color=species))
```

```{r}
ggplot(data=penguins, aes(x=bill_length_mm, y=bill_depth_mm))+
  geom_point(aes(color=species))+
  geom_smooth(aes(color=species))
```
```{r}
penguins_copy <- data.frame(penguins)

penguins_copy <- penguins_copy  %>% 
  mutate(bill_depth_mm=replace(bill_depth_mm, bill_length_mm==58.0, 22))
# 
# penguins_copy %>% 
#   filter(species=='Chinstrap') %>% 
#   arrange(-bill_length_mm)

ggplot(data=penguins_copy, aes(x=bill_length_mm, y=bill_depth_mm))+
  geom_point(aes(color=species))+
  geom_smooth(aes(color=species))

```


```{r}
penguins %>% 
  filter(species=='Chinstrap') %>% 
  arrange(-bill_length_mm)
```

in the above output the **Chinstrap** specie with the **bill_length_mm=58.0** seems to be an outlier for it or just might be classified wrongly as it should be a **Gentoo** specie.
which can be shown in the smooth_line drawn in the above figure.
or it might be a conversion as the bill_length_mm increased the penguin started to converts from **Chinstrap** to **Gentoo** .

```{r}
ggplot(data=penguins) +
  geom_histogram(mapping=aes(x=flipper_length_mm, fill=species), bins = 30)
```

```{r}
ggplot(data=penguins) +
  geom_histogram(mapping=aes(y=flipper_length_mm, fill=species), bins = 30)+
  facet_wrap(~species)
```

```{r}
ggplot(penguins)+
  geom_bar(aes(sex, fill=sex))
```

```{r}
ggplot(penguins)+
  geom_bar(aes(island, fill=island))
```

```{r}
ggplot(penguins)+
  geom_point(aes(x=flipper_length_mm, y=body_mass_g,
                 color=species, shape=sex), alpha=.7)+
  facet_grid(island~species)+
  labs(title="Palmer Penguins: Flipper Length vs. Body Mass",
       subtitle='Data grouped with island and species')
```

As shown in the above plot. We found that **Adelie** specie found in all the three island that why it has more records. in the dataset and the **Chinstrap** specie only shown in the **Dream** island and the **Gentoo** specie shown only in the **Biscoe** island.